---
title: "Civic Accountability Scorecard"
subtitle: "`r params$project`"
author: "`r params$analyst`"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M UTC', tz='UTC')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
    self_contained: true
    css: civic_report_style.css
params:
  object:     NULL     # civic_model
  test_data:  NULL     # data.frame
  outcome:    "outcome"
  protected:  NULL
  positive:   NULL
  project:    "civic.icarm"
  analyst:    "Anonymous"
  notes:      ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 8, fig.height = 4.5, dpi = 120
)
library(civic.icarm)
library(ggplot2)
library(dplyr)
library(knitr)

object    <- params$object
test_data <- params$test_data
outcome   <- params$outcome
protected <- params$protected
positive  <- params$positive
```

---

## 1. Model Provenance

```{r provenance}
prov <- data.frame(
  Field = c("Model type", "Learner", "Formula", "N train",
            "Seed", "Trained at", "Data hash (SHA-256, first 24 chars)"),
  Value = c(
    object$type,
    object$model,
    deparse(object$formula),
    format(object$n_train, big.mark = ","),
    as.character(object$seed),
    format(object$trained_at, "%Y-%m-%d %H:%M:%S UTC", tz = "UTC"),
    substr(object$data_hash, 1, 24)
  )
)
kable(prov, col.names = c("Field", "Value"), align = c("l","l"))
```

> **DataCitizen-Pro — Data Literacy pillar:** Full provenance enables any
> analyst to reproduce this exact model. The data hash detects tampering.

---

## 2. Test Set Performance

```{r performance}
y_true <- test_data[[outcome]]

if (object$type == "classification") {
  y_hat  <- predict(object, test_data, type = "class")
  y_prob <- tryCatch({
    p <- predict(object, test_data, type = "prob")
    if (is.matrix(p)) p[, ncol(p)] else as.numeric(p)
  }, error = function(e) NULL)
  perf <- civic_metrics(y_true, y_hat, y_prob = y_prob, positive = positive)
} else {
  y_hat <- predict(object, test_data)
  perf  <- civic_metrics(y_true, y_hat, type = "regression")
}

perf_df <- data.frame(
  Metric = names(perf),
  Value  = round(as.numeric(perf), 4)
)
kable(perf_df, align = c("l","r"))
```

```{r perf-plot, fig.height=3}
if (object$type == "classification") {
  civic_plot_confusion(y_true, y_hat)
}
```

> **DataCitizen-Pro — Statistical Reasoning pillar:** Performance metrics
> must be interpreted in context. Accuracy alone is insufficient; balanced
> accuracy accounts for class imbalance.

---

## 3. Model Explanation

```{r explain}
ex <- civic_explain(object)
```

```{r importance-plot}
if (!is.null(ex$importance)) {
  civic_plot_importance(ex)
}
```

**Importance method:** `r ex$importance_method`

```{r importance-table}
if (!is.null(ex$importance)) {
  kable(head(ex$importance[, c("feature","importance","importance_scaled")], 10),
        digits = 4, align = c("l","r","r"))
}
```

---

## 4. Probability Calibration

```{r calibration, fig.height=4}
if (object$type == "classification" && !is.null(y_prob)) {
  cal <- tryCatch(
    civic_calibration(object, test_data, outcome, positive),
    error = function(e) NULL
  )
  if (!is.null(cal)) {
    print(cal)
    civic_plot_calibration(cal)
  }
}
```

> **DataCitizen-Pro — Statistical Reasoning pillar:** A model used to
> communicate risk to citizens must be well-calibrated. An ECE > 0.10
> indicates systematic over- or under-confidence.

---

## 5. Fairness Audit

```{r fairness-section, results='asis'}
if (is.null(protected)) {
  cat("> **No protected attribute specified.** Fairness audit not conducted.\n\n")
  cat("> Re-render this report with `protected = 'column_name'` to include equity metrics.\n")
}
```

```{r fairness, eval=!is.null(protected)}
fair <- civic_fairness_report(object, test_data, outcome, protected, positive)
kable(round(dplyr::select(fair, where(is.numeric)), 3))
```

```{r fairness-tpr, eval=!is.null(protected), fig.height=3.5}
civic_plot_fairness(fair, metric = "tpr",
                    title = "Equal Opportunity: TPR by Group")
```

```{r fairness-dp, eval=!is.null(protected), fig.height=3.5}
civic_plot_fairness(fair, metric = "dp_ratio",
                    title = "Disparate Impact Ratio",
                    ref_line = 0.8)
```

```{r fairness-eo, eval=!is.null(protected), fig.height=4.5}
tryCatch({
  eoc <- civic_equalized_odds_curve(object, test_data, outcome, protected, positive)
  civic_plot_roc_groups(eoc)
}, error = function(e) NULL)
```

```{r equity-summary, eval=!is.null(protected)}
eq <- civic_equity_summary(fair)
eq_df <- data.frame(
  Indicator = names(eq),
  Value = sapply(eq, function(v) {
    if (is.logical(v)) ifelse(v, "PASS ✓", "FAIL ✗") else round(as.numeric(v), 4)
  })
)
kable(eq_df, align = c("l","r"))
```

> **DataCitizen-Pro — Democratic Judgment pillar:** The EU AI Act (2024)
> requires fairness reporting for high-risk algorithmic systems. Discuss
> with learners: *which fairness criterion is most appropriate here, and why?*

---

## 6. Threshold Analysis

```{r threshold-plot, eval=object$type=="classification" && !is.null(y_prob)}
thr_tbl <- civic_thresholds(y_true, y_prob, positive = positive)
civic_plot_thresholds(thr_tbl,
  metrics = c("accuracy", "recall", "precision", "f1"))
```

> **DataCitizen-Pro:** The choice of classification threshold is a *democratic*
> decision. Who bears the cost of false positives vs false negatives?

---

## 7. Audit Trail

```{r audit-trail}
trail <- civic_audit_trail(
  object   = object,
  metrics  = perf,
  fairness = if (exists("fair")) fair else NULL,
  analyst  = params$analyst,
  notes    = params$notes
)
cat("```json\n")
cat(trail)
cat("\n```\n")
```

---

## 8. Analyst Notes

`r if (nchar(params$notes) > 0) params$notes else "*No notes provided.*"`

---

## 9. Session Info

```{r session, echo=TRUE}
sessionInfo()
```

---

*Generated by [civic.icarm](https://github.com/oluwale-awe/civic.icarm) —
DataCitizen-Pro / Ludwigsburg University of Education.*
